---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading libraries 
```{r echo=TRUE}
library(dplyr)
library(ggplot2)
```

## Loading and preprocessing the data
```{r echo=TRUE}
urlData <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
zipData <- "activity.zip"

fileData <- "activity.csv"
dateFormat <- "%Y-%m-%d"

weekenddays <- c("sáb", "dom.") #Spanish shortnames for saturday and sunday

if (!file.exists(fileData)) { 
  if (!file.exists(zipData)){
    download.file(urlData, zipData)
  }
  unzip(zipData)
}

setAs("character","sdate", function(from) as.Date(from, format=dateFormat) )
activity <- read.csv(fileData, colClasses = c("numeric", "sdate", "numeric"))
```
## What is mean total number of steps taken per day?
```{r echo=TRUE}
activityperday <- activity %>% na.omit() %>% group_by(date) %>% summarise(steps = sum(steps))
ggplot(activityperday, aes(x = steps)) + geom_histogram()

ggplot(activityperday, aes(x = date, y = steps)) + geom_bar(stat = "identity")

mean(activityperday$steps)
median(activityperday$steps)

```

## What is the average daily activity pattern?
```{r}
activityperinterval <- activity %>% na.omit() %>% group_by(interval) %>% summarise(steps = mean(steps))
ggplot(activityperinterval, aes(x=interval, y=steps)) + geom_line()

maxsteps <- max(activityperinterval$steps)
activityperinterval[activityperinterval$steps == maxsteps,]

```



## Imputing missing values
```{r}
sum(is.na(activity$steps))

stepsintervalmean <- mean(activity$steps, na.rm = TRUE)

activitywona <- activity
activitywona[is.na(activitywona$steps),]$steps <- stepsintervalmean

activitywonaperday <- activitywona %>% na.omit() %>% group_by(date) %>% summarise(steps = sum(steps))

ggplot(activitywonaperday, aes(x = steps)) + geom_histogram()

mean(activitywonaperday$steps)
median(activitywonaperday$steps)

```
Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?


## Are there differences in activity patterns between weekdays and weekends?
```{r}
activitywona$typeday <- NA
activitywona[weekdays(activitywona$date, abbreviate =TRUE) %in% weekenddays,]$typeday <- "weekend"

activitywona[is.na(activitywona$typeday),]$typeday <- "weekday"

activitywonaperinterval <- activitywona %>% group_by(typeday,interval) %>% summarise(steps = mean(steps))
ggplot(activitywonaperinterval, aes(x=interval, y=steps)) + geom_line() + facet_grid(typeday~.)


```

